---
title: "STATS 100 Investigations"
output: 
  learnr::tutorial:
    progressive: FALSE
    allow_skip: FALSE
    df_print: paged
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(lubridate)
library(stringr)
library(glue)
library(rvest)
library(jsonlite)
library(knitr)
library(httr)

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)
tutorial_options(exercise.timelimit = 90)

# I think the stuff below means the previous code is not cached but ???
options(tutorial.storage = list(
  # save an arbitrary R object "data" to storage
   save_object = function(tutorial_id, tutorial_version, user_id, object_id, data) {},
  # retreive a single R object from storage
  get_object = function(tutorial_id, tutorial_version, user_id, object_id) { 
    NULL 
  },
  
  # retreive a list of all R objects stored
  get_objects = function(tutorial_id, tutorial_version, user_id) { 
    list() 
  },
  
  # remove all stored R objects
  remove_all_objects = function(tutorial_id, tutorial_version, user_id) {
  }
))

getMovieData <- function(keyword, password, data_mode){
  # get API key
  url <- glue("https://script.google.com/macros/s/AKfycbynYqGYeHLNDjcltlU5jGHx1ZPWWvqbelq61tj8eAaZC71ByTA/exec?password={password}")
  response <- httr::GET(url = url)
  api_key <- httr::content(response, "text")
  if(api_key != "")
  {
    if(data_mode == "testing")
    {
      pages <- c(2,4,6,8,10)
          return(searchMovieData(keyword, api_key, pages))
    }
     if(data_mode == "training")
    {
      pages <- c(1,3,5,7,9)
          return(searchMovieData(keyword, api_key, pages))
    }
  }
  else
  {
    return()
  }
}

testModel <- function(data, a, b, error){
    if(nrow(data) > 0)
    {
      points_plotted <- data %>%
        filter(!Metascore == "N/A") %>%
        nrow()

      data %>%
        filter(!is.na(Metascore), 
               !is.na(imdbRating)) %>%
        mutate(Metascore = Metascore %>% as.numeric(),
               imdbRating = imdbRating %>% as.numeric()) %>%
        ggplot(aes(x = imdbRating, 
                   y = Metascore)) + 
        geom_point(alpha = 0.5) + 
        geom_abline(intercept = a, 
                    slope = b) +  
        geom_ribbon(aes(ymin = a + b*imdbRating - error, 
                        ymax = a + b*imdbRating  + error), 
                    fill = "#f1c232", 
                    alpha = 0.5) + 
        annotate("text", -Inf, Inf, 
                 label = paste0("n =",
                                points_plotted), 
                 hjust = 0, 
                 vjust = 1)  +
theme_minimal()
    }
}

searchMovieData <- function(keyword, api_key, pages){
  keyword <- keyword %>% str_trim() %>% str_replace_all(" ","+")
  url <- glue("http://www.omdbapi.com/?s={keyword}&type=movie&apikey={api_key}")
  results <- fromJSON(url)
  numResults <- results$totalResults %>% as.numeric()
  if(numResults < 100){
    return()
  }
  else
  {
    
    # get all the IDs to look up
    search_data <- map_df(1 : length(pages), function(i){
      page_num <- pages[i]
      url <- glue("http://www.omdbapi.com/?s={keyword}&type=movie&apikey={api_key}&page={page_num}")
      results <- fromJSON(url)
      results$Search
    })
    
    # look up each movie
    movie_ids <- search_data$imdbID
    movie_data <- map_df(1 : length(movie_ids), function(i){
        id <- movie_ids[i]
        url <- glue("http://www.omdbapi.com/?i={id}&apikey={api_key}")
        response <- fromJSON(url)
        tibble(Title = response$Title,
               Year = response$Year,
               Metascore = response$Metascore,
               imdbRating = response$imdbRating)
    })
    return(movie_data)
  }
}

# function to get data out and into iNZightLite
iNZightLink <- function(data) {
  json_output <- jsonlite::toJSON(data, na = "null")
  anna_url <- "https://csv.monster/tempcsv.php"
  response <- httr::POST(url = anna_url, body = list(t = json_output))
  temp_csv_url <- httr::content(response, "text")
  src <- glue("https://lite.docker.stat.auckland.ac.nz/?url={temp_csv_url}&land=visualize")
  data.frame(link_to_iNZight_lite = glue("<a href='{src}' target = '_blank'>{src}</a>")) %>% kable()
}

```

```{css}

@import url('https://fonts.googleapis.com/css?family=Inconsolata|Londrina+Solid&display=swap');

.tutorialTitle:before {
  content: url("images/STATS100-Cover-Logo_small.png");
  display: block;
}

h1, h2, h3{
  font-family: 'Londrina Solid', cursive;
}

h3 {
  color: #f5af86;
}

h4 {
  font-family: 'Inconsolata', monospace;
  font-weight: bold;
}

.btn-primary {
  background-color: #f5af86;
}

.btn-primary:hover, .btn-primary:focus {
  background-color: #f5af86;
  opacity: 0.8;
}

.myButtonSelected {
  background-color:#666666;
}

.myButtonSelected:hover, .myButtonSelected:focus{
  background-color:#666666;
}

body, .btn{
  font-family: 'Inconsolata', monospace;
  font-size: 16px;
}

a {
  color: #000000;
  font-weight: bold;
  background-color: #F0F0F0;
}

a:hover, a:focus {
  color: #000000;
  opacity: 0.8;
}

blockquote {
  font-size: 16px;
  background-color: #f5af86;
  padding: 10px;
  color: #000000;
}

.highlight {
  background-color: #F0F0F0;
}

td {
  padding: 10px;
}

tr {
  border: solid #ddd 1px;
}

.center-align{ text-align:center;}
```

```{js echo=FALSE}
$(function() {
  var editor;
  $('.ace_editor').each(function( index ) {
    editor = ace.edit(this);
    editor.getSession().setUseWrapMode(true);
    editor.setFontSize("16px");
    });

  $(".topic").click(function(){
    if($(this).text() == "Return to home page"){
    window.location.href = "https://stats100.docker.stat.auckland.ac.nz/";
    }
  })
  
})
```

## Movie ratings

Can we use a movie's metascore to predict its IMDb rating? 

We'll use data sourced from the <a href="http://www.omdbapi.com/" target="blank">OMDb API</a> to investigate.

### Training the model

 + Change the keyword to your keyword, and enter the password between the "". 
 + Then run the code and click the link generated to explore the data using iNZight Lite.

> This code will take a while to run!

```{r trainModel, exercise=TRUE}
# get training data
keyword <- "cat"
password <- ""
data_mode <- "training"
data <- getMovieData(keyword, password, data_mode)

# visualise & model
iNZightLink(data)
```

### Testing the model

 + Change the keyword to your keyword, and enter the password between the "". 
 + Change the model parameters to your values!
 + Then run the code to see how well your model performs on new data!

> This code will take a while to run!

```{r testModel, exercise=TRUE}
# get testing data
keyword <- "cat"
password <- ""
data_mode <- "testing"
data <- getMovieData(keyword, password, data_mode)

# test model
a <- 0
b <- 0
error <- 0
testModel(data, a, b, error)
```

## <a>Return to home page</a>