---
title: "STATS 100 Investigations"
output: 
  learnr::tutorial:
    progressive: FALSE
    allow_skip: FALSE
    df_print: paged
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(stringr)
library(glue)
library(knitr)
library(httr)
library(spotifyr)

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)
tutorial_options(exercise.timelimit = 90)

# clustering function to reuse
getSongs <- function(artist_name, password, k = 0){
 url <- glue("https://script.google.com/a/aucklanduni.ac.nz/macros/s/AKfycbz4Qme0IZxDn6i_ezuq0DQ-azy0CH3vdPfXs9Ll/exec?password={password}")
  response <- httr::GET(url = url)
  api_key <- httr::content(response, "text")
  if(api_key != ""){
      api_info <- api_key %>% str_split(",")
      Sys.setenv(SPOTIFY_CLIENT_ID = api_info[[1]][1])
      Sys.setenv(SPOTIFY_CLIENT_SECRET = api_info[[1]][2])
      access_token <- get_spotify_access_token()
      songs <- get_artist_audio_features(artist_name)
      songs_tidy <- songs %>%
        mutate(mode_num = ifelse(mode_name == "minor", 0, 1)) %>%
        group_by(artist_name, 
                 str_to_lower(track_name)) %>%
        summarise(energy = max(energy),
                valence = max(valence),
                mode_num = mean(mode_num) %>% round(),
                track_uri = first(track_uri),
                album_name = first(album_name)) %>%
        mutate(mode_name = ifelse(mode_num == 1, "major", "minor")) %>%
        select(-mode_num) %>%
        rename(track_name = `str_to_lower(track_name)`)

      songs_features <- songs_tidy %>%
        ungroup() %>%
        select(energy, valence)

      # cluster time!
      set.seed(2020)
      if(k == 0)
      {
        for(i in 2 : round(nrow(songs_tidy)/5)){
                num_clust <- i
                songs_km <- kmeans(songs_features, 
                           centers = num_clust) 
                new_var <- glue("cluster_k_{i}")
                songs_tidy[[new_var]] <- as.factor(songs_km$cluster %>% paste("cluster", .))
        }
      }
      else
      {
        num_clust <- k
        songs_km <- kmeans(songs_features, 
                           centers = num_clust) 
        songs_tidy$cluster <- as.factor(songs_km$cluster)
      }
      return(songs_tidy)
    }
    else
    {
      stop("That password is not correct")
    } 
}

# function to get song data
getSongData <- function(artist_name, password){
  if(password == ""){
    stop("Did you forget to enter the password?")
  }
  else
  {
    getSongs(artist_name, password)
  }
}

# function to return two songs for each cluster, using spotify web player embedded iframes
getSongClusters <- function(artist_name, password, k){
  if(password == ""){
    stop("Did you forget to enter the password?")
  }
  else
  {
    if(k < 2)
    {
       stop("You need to provide a value for k that is at least 2 (e.g. at least 2 clusters/groups)")
    }
    else
    {
      cluster_songs <- getSongs(artist_name, password, k)
      # try one for now
      track_data <- cluster_songs %>%
        mutate(iframe_src = str_replace_all(track_uri, "spotify:track:","")) %>%
        arrange(cluster) %>%
        mutate(player = paste0("<iframe src='https://open.spotify.com/embed/track/", iframe_src, "' width='500' height='100' frameborder='0' allowtransparency='true' allow='encrypted-media'></iframe>")) %>%
        select(track_name, cluster, player) %>%
        group_by(cluster) %>% 
        sample_n(min(3, n()))
    
      track_data %>% kable()
    }
  }
}

######  SHOW THE PLAYER FOR THE TRACKS
#<iframe src="https://open.spotify.com/embed/track/4aebBr4JAihzJQR0CiIZJv" width="300" height="380" frameborder="0" allowtransparency="true" allow="encrypted-media"></iframe>
######

# function to get data out and into iNZightLite
iNZightLink <- function(data) {
  json_output <- jsonlite::toJSON(data, na = "null")
  anna_url <- "https://csv.monster/tempcsv.php"
  response <- httr::POST(url = anna_url, body = list(t = json_output))
  temp_csv_url <- httr::content(response, "text")
  src <- glue("https://lite.docker.stat.auckland.ac.nz/?url={temp_csv_url}&land=visualize")
  data.frame(link_to_iNZight_lite = glue("<a href='{src}' target = '_blank'>{src}</a>")) %>% kable()
}

```

```{css}

@import url('https://fonts.googleapis.com/css?family=Inconsolata|Londrina+Solid&display=swap');

.tutorialTitle:before {
  content: url("images/STATS100-Cover-Logo_small.png");
  display: block;
}

h1, h2, h3{
  font-family: 'Londrina Solid', cursive;
}

h3 {
  color: #f5af86;
}

h4 {
  font-family: 'Inconsolata', monospace;
  font-weight: bold;
}

.btn-primary {
  background-color: #f5af86;
}

.btn-primary:hover, .btn-primary:focus {
  background-color: #f5af86;
  opacity: 0.8;
}

.myButtonSelected {
  background-color:#666666;
}

.myButtonSelected:hover, .myButtonSelected:focus{
  background-color:#666666;
}

body, .btn{
  font-family: 'Inconsolata', monospace;
  font-size: 16px;
}

a {
  color: #000000;
  font-weight: bold;
  background-color: #F0F0F0;
}

a:hover, a:focus {
  color: #000000;
  opacity: 0.8;
}

blockquote {
  font-size: 16px;
  background-color: #f5af86;
  padding: 10px;
  color: #000000;
}

.highlight {
  background-color: #F0F0F0;
}

td {
  padding: 10px;
}

tr {
  border: solid #ddd 1px;
}

.center-align{ text-align:center;}
```

```{js echo=FALSE}
$(function() {
  var editor;
  $('.ace_editor').each(function( index ) {
    editor = ace.edit(this);
    editor.getSession().setUseWrapMode(true);
    editor.setFontSize("16px");
    });

  $(".topic").click(function(){
    if($(this).text() == "Return to home page"){
    window.location.href = "https://stats100.docker.stat.auckland.ac.nz/";
    }
  })
  
})
```

## Song sorter

Can we sort songs from an artist into similar sounding songs based on two measures: energy and valence? 

We'll use data sourced from the <a href="https://developer.spotify.com/documentation/web-api/reference/tracks/get-audio-features/" target="blank">Spotify API</a> to investigate using *k-means clustering*.

### Exploring the data and possible clusters

 + Change the artist name to your artist, and **enter the password between the ""**. 
 + Then run the code and click the link generated to explore the data using iNZight Lite.

> This code will take a while to run!

```{r trainModel, exercise=TRUE}
# get song data based on artist
artist_name <- "adele"
password <- ""

data <- getSongData(artist_name, password)

# visualise & model
iNZightLink(data)
```

### But do the songs sound similar?

 + Change the artist name to your artist, and enter the password between the "". 
 + Enter the number of clusters you decided on for your songs (k)
 + Then run the code and listen to a random selection of up to three of the songs from each of these clusters
 
> This code will take a while to run!
 
```{r testModel, exercise=TRUE, exercise.lines = 7}
# get song cluster data based on artist and k
artist_name <- "adele"
password <- ""
k <- 0

getSongClusters(artist_name, password, k)
```

## <a>Return to home page</a>