---
title: "STATS 100 Investigations"
output: 
  learnr::tutorial:
    progressive: TRUE
    allow_skip: FALSE
    df_print: paged
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(lubridate)
library(stringr)
library(glue)
library(rvest)
library(jsonlite)
library(knitr)
library(httr)

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)
tutorial_options(exercise.timelimit = 90)

# I think the stuff below means the previous code is not cached but ???
options(tutorial.storage = list(
  # save an arbitrary R object "data" to storage
   save_object = function(tutorial_id, tutorial_version, user_id, object_id, data) {},
  # retreive a single R object from storage
  get_object = function(tutorial_id, tutorial_version, user_id, object_id) { 
    NULL 
  },
  
  # retreive a list of all R objects stored
  get_objects = function(tutorial_id, tutorial_version, user_id) { 
    list() 
  },
  
  # remove all stored R objects
  remove_all_objects = function(tutorial_id, tutorial_version, user_id) {
  }
))

# some functions
getMusicData <- function(chart_id){
  week_url <- glue("https://nztop40.co.nz/chart/singles?chart={chart_id}")
    week_page <- read_html(week_url)
    
    week_date <- week_page %>%
      html_nodes("#p_calendar_heading") %>%
      html_text() %>%
      unique()
    
    single_position <- week_page %>%
      html_nodes("td.p_this_week > p > span") %>%
      html_text() %>%
      str_squish() %>%
      head(40)
    
    single_last_week <- week_page %>%
      html_nodes("td.p_last_week") %>%
      html_text() %>%
      str_squish() %>%
      head(40)
    
    single_weeks <- week_page %>%
      html_nodes("td.p_weeks") %>%
      html_text() %>%
      str_squish() %>%
      head(40)
    
    single_title <- week_page %>%
      html_nodes("td.p_title_artist > p.p_title") %>%
      html_text() %>%
      str_squish() %>%
      head(40)
    
    single_artist <- week_page %>%
      html_nodes("td.p_title_artist > p.p_artist") %>%
      html_text() %>%
      str_squish() %>%
      head(40)
    
    nztop40 <- tibble(single_position, 
            single_last_week,
            single_weeks,
            single_title,
            single_artist) %>%
      mutate(week_date)
    
    # add info from itunes
    track_apple_info <- map_df(1 : nrow(nztop40), function(i){
      # wait 30 seconds
      Sys.sleep(0.5)
      search_artist <- nztop40$single_artist[i]
      search_title <- nztop40$single_title[i]
      track_info <- getTrackDetails(search_artist, search_title)
      
      if(!is.null(track_info)){
        info <- track_info %>%
          select(trackId, artistName, collectionName, trackName, releaseDate, trackTimeMillis, primaryGenreName) %>%
          mutate(single_artist = search_artist, 
                 single_title = search_title)
        
        if(nrow(info) > 0 && !is.null(info)){
          return(info)
        }
        
      }
  })
    
    nztop40 %>%
      left_join(track_apple_info) %>%
      mutate(single_last_week = ifelse(str_detect(single_last_week, "new|re-entry"),
                                   "",
                                   single_last_week),
             apple_track_length_mins = trackTimeMillis/60000,
             apple_genre = fct_lump_n(primaryGenreName, 2),
             apple_days_since_released = difftime(
               parse_date_time(
                 week_date, 
                 orders="dmy"),
               parse_date_time(
                 releaseDate %>% 
                   str_sub(1,10), 
                 orders="ymd"),
               units="days") %>% 
               as.numeric(),
             apple_days_since_released = ifelse(
               apple_days_since_released < 0, 
               0, apple_days_since_released),
             single_title_num_words = str_count(single_title,
                                                "\\S+"),
             single_artist_featured = ifelse(
               str_detect(single_artist, "feat."),
               "yes",
               "no"),
             single_moved_position = ifelse(
               single_position != single_last_week,
               "moved chart position",
               "same chart position or new")
             ) %>%
      rename(apple_id = trackId,
             apple_artist = artistName,
             apple_title = trackName,
             apple_collection = collectionName,
             apple_release_date = releaseDate,
             chart_date = week_date) %>%
      select(-trackTimeMillis, 
             -primaryGenreName, 
             chart_date,
             starts_with("single"),
             starts_with("apple"))
}

#a <- parse_date_time("20 April 2020", orders="dmy")
#b <- parse_date_time("2018-11-08T08:00:00Z" %>% str_sub(1,10), orders="ymd")
#difftime(parse_date_time(week_date, orders="dmy"),parse_date_time(releaseDate %>% str_sub(1,10), orders="ymd"),units="days") %>% as.numeric()

#get info from itunes
getTrackDetails <- function(artist, track){
  search_artist <- artist %>% str_replace_all(" ","+")
  # need to remove feat. Drake etc.???
  apple_url <- glue("https://itunes.apple.com/search?term={search_artist}")
  artist_song <- fromJSON(apple_url, flatten = TRUE)$results 
  if(!length(artist_song %>% unlist()) == 0)
  {
   return(artist_song %>%
             filter(str_detect(trackName %>% str_to_lower() %>% str_replace_all("[^[:alnum:]]", ""), 
                               track %>% str_to_lower() %>% str_replace_all("[^[:alnum:]]", "") %>% str_replace_all("\\(.*\\)", "") %>% str_squish())) %>%
             head(1))
  }
}

# function to get data out and into iNZightLite
iNZightLink <- function(data) {
  json_output <- jsonlite::toJSON(data, na = "null")
  anna_url <- "https://csv.monster/tempcsv.php"
  response <- httr::POST(url = anna_url, body = list(t = json_output))
  temp_csv_url <- httr::content(response, "text")
  src <- glue("https://lite.docker.stat.auckland.ac.nz/?url={temp_csv_url}&land=visualize")
  data.frame(link_to_iNZight_lite = glue("<a href='{src}' target = '_blank'>{src}</a>")) %>% kable()
}

```

```{css}

@import url('https://fonts.googleapis.com/css?family=Inconsolata|Londrina+Solid&display=swap');

.tutorialTitle:before {
  content: url("images/STATS100-Cover-Logo_small.png");
  display: block;
}

h1, h2, h3{
  font-family: 'Londrina Solid', cursive;
}

h3 {
  color: #f5af86;
}

h4 {
  font-family: 'Inconsolata', monospace;
  font-weight: bold;
}

.btn-primary {
  background-color: #f5af86;
}

.btn-primary:hover, .btn-primary:focus {
  background-color: #f5af86;
  opacity: 0.8;
}

.myButtonSelected {
  background-color:#666666;
}

.myButtonSelected:hover, .myButtonSelected:focus{
  background-color:#666666;
}

body, .btn{
  font-family: 'Inconsolata', monospace;
  font-size: 16px;
}

a {
  color: #000000;
  font-weight: bold;
  background-color: #F0F0F0;
}

a:hover, a:focus {
  color: #000000;
  opacity: 0.8;
}

blockquote {
  font-size: 16px;
  background-color: #f5af86;
  padding: 10px;
  color: #000000;
}

.highlight {
  background-color: #F0F0F0;
}

td {
  padding: 10px;
}

tr {
  border: solid #ddd 1px;
}

.center-align{ text-align:center;}
```

```{js echo=FALSE}
$(function() {
  var editor;
  $('.ace_editor').each(function( index ) {
    editor = ace.edit(this);
    editor.getSession().setUseWrapMode(true);
    editor.setFontSize("16px");
    });



  $(".topic").click(function(){
    if($(this).text() == "Return to home page"){
    window.location.href = "https://stats100.docker.stat.auckland.ac.nz/";
    }
  })
  
})
```

```{r eval=FALSE}
parse_date("20 April 2020")

release_date <- "2020-02-01T08:00:00Z" %>%
  str_replace("T.*Z", "") %>%
  parse_date()

data <- getMusicData(chart_id = 4848)
iNZightLink(data)


```

## Birthday songs

What songs were in the NZ Top 40 music charts the week you were born? 

Change the `chart id` number in the code below to match the number from the URL from the week you were born. 

The `chart_id` is the number after **?chart=** in the URL `https://nztop40.co.nz/singles?chart=ID`

Then run the code and explore the data about these songs using iNZight Lite!

> Be patient, this code will take about 30 seconds to run!

```{r getData, exercise=TRUE}
data <- getMusicData(chart_id = 4848)
iNZightLink(data)
```

## <a>Return to home page</a>